---
title: "Assignment 06"
author: "Noah Scheider"
date: "11/18/2022"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r data, include=FALSE}
library(VGAM)
```

# Exercise

We currently have $3$ ways of generating Gaussian samples in R:

* the native `rnorm()` using the inverse transform 
* Box-Muller transform
* rejection sampling (see lecture notes for code)

For $50$ simulation runs, and for all three methods:

* generate 105 samples using
* time the generation process using `Sys.time()`
* calculate the KS statistic ([the one-sample Kolmogorov-Smirnov statistic](https://en.wikipedia.org/wiki/Kolmogorov%E2%80%93Smirnov_test)) to check quality of the sample, e.g.
  
  * `X <- rnorm(10ˆ5)`
  * `F_hat <- ecdf(X)`
  * `F_hat_eval <- F_hat(sort(X))`
  * `F_eval <- pnorm(sort(X))`
  * `KS_stats <- max(abs(F_hat_eval - F_eval))`

Compare the three methods using your simulation results.

```{r fig.align='center'}

```

We see that

