---
title: "Assignment 04"
author: "Noah Scheider"
date: "11/04/2022"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(MASS)
library(locpol)
```

# Exercise

Consider a subset of mcycle data (of the MASS package) for times $≤ 40$ and use cross-validation to select

* the polynomial degree of p from candidate values $p = 1, 2, 3$, and 
* the bandwidth $h$ from candidate values $h = 3, 4, . . . , 15$

for a local polynomial smoother as implemented by the `locpol()` function from the package of the same name.

Notes:

* Compare your results with what you would expect based on Manual 10 in order to avoid wrong conclusions.
  * even better: use your own visualizations to verify your progress
* You may run into issues for large p and small h if you use small number of folds.
* Beware of how your points are ordered.

```{r fig.cap=c("Figure that plots the different degrees p of a local polynomial regression against different bandwidths (from 3:15)", "plots the local polynomial regression with the bandwidth and the degree with the smallest MSE"),  fig.align='center', echo=FALSE}
data(mcycle)
mcycle <- mcycle[mcycle$times <= 40, ]
p_degree <- c(1, 2, 3)
hgrid <- 3:15

# data frame to capture different errors
df_err <- data.frame(matrix(nrow=length(hgrid), ncol=3))
colnames(df_err) <- c("Degree 1", "Degree 2", "Degree 3")
rownames(df_err) <- c(hgrid)

# length of data set
n <- length(mcycle$times)

# LOO and MSE
for (p in p_degree) {
  for (h in hgrid) {
    mse <- 0
    for (i in 1:n) {
      model <- locpol(formula = accel[-i]~times[-i], data = mcycle, bw = h, deg = p, xeval = mcycle$times[i])
      mse <- mse + (mcycle$accel[i]-model$lpFit[2])^2/n
    }
    df_err[which(rownames(df_err) == h), p] <- mse
  }
}

plot(hgrid, df_err$`Degree 1`, type = 'l', col = 1, ylab = "MSE")
lines(hgrid, df_err$`Degree 2`, col = 2)
lines(hgrid, df_err$`Degree 3`, col = 3)
legend('topleft', legend = c(colnames(df_err)), col = 1:3, pch = 16)

min_min <- which(df_err == min(df_err, na.rm = TRUE), arr.ind = TRUE)
best_model <- locpol(formula = accel~times, data = mcycle, bw = min_min[1], deg = min_min[2], xeval = mcycle$times)
plot(mcycle$times, mcycle$accel)
lines(mcycle$times, best_model$lpFit$accel)
legend('topleft', legend = sprintf("h = %.f, p = %.f, MSE = %.2f", min_min[1], min_min[2], df_err[min_min]))

```


