---
title: "Assignment_03"
author: "Noah Scheider"
date: "October 14, 2022"
header-includes: \usepackage{bbm}
output: bookdown::html_document2
---

```{r setup, echo=FALSE}
library(bookdown)
```

# Exercise

Slide “Bandwidth > Kernel” above gives a vague statement, that choosing bandwidth h is more important than choosing a proper kernel. See this for yourself using a small simulation study. Specifically:

- use Manual 09 to generate data from the Gaussian mixture $f$ 
- repeat the following 200 times:
  1. Generate $N = 100$ samples from the Gaussian mixture
  2. Perform density estimation, i.e. obtain $\hat f$ for Gaussian, Epanechnikov, and rectangular kernels as well as bandwidth values $h = 0.1, 0.15, 0.2, 0.25, . . . , 0.9$
  3. Calculate the error measure $||f-\hat f||_{2}$
- report your findings as a single (well commented) figure

**Note**: Please check again Section 5 of the Course Organization for submission instructions.


```{r, echo=FALSE}
# rmixnorm and dmixnorm imported from the Module 09_GaussianMixture.html

# returns RV for a Gaussian mixture
rmixnorm <- function(N, mu1, mu2, sigma1, sigma2, tau){
  ind <- I(runif(N) > tau)
  X <- rep(0,N)
  X[ind] <- rnorm(sum(ind), mu1, sigma1)
  X[!ind] <- rnorm(sum(!ind), mu2, sigma2)
  return(X)
}

# theoretical pdf for a Gaussian mixture
dmixnorm <- function(x, mu1, mu2, sigma1, sigma2, tau){
  y <- (1-tau)*dnorm(x,mu1,sigma1) + tau*dnorm(x,mu2,sigma2)
  return(y)
}

n = 200

bw = seq(0.1, 0.9, 0.05)

gau_err <- matrix(nrow = n, ncol = length(bw))
epane_err <- matrix(nrow = n, ncol = length(bw))
rect_err <- matrix(nrow = n, ncol = length(bw))

for (i in 1:n) {
  t = 0.25
  # Generate 100 samples from the Gaussian mixture
  rv <- rmixnorm(100, 0, 0, 1, 1, t)
  # density estimation for different Kernels with Bandwiths 0.1, 0.15, 0.2
  for (k in 1:length(bw)){
    # error of Gaussian Kernel
    gaussian <- density(rv, kernel="gaussian", bw=bw[k])
    gau_err[i, k] <- norm(sapply(gaussian$x, function(z) dmixnorm(z, 0, 0, 1, 1, t))-gaussian$y, type = c("2"))
    # error of Epanechnikov Kernel
    epane <-density(rv, kernel="epanechnikov", bw=bw[k])
    epane_err[i, k] <- norm(sapply(epane$x, function(z) dmixnorm(z, 0, 0, 1, 1, t))-epane$y, type = c("2"))
    # error of Rectangular Kernel
    rect <- density(rv, kernel="rectangular", bw=bw[k])
    rect_err[i, k] <- norm(sapply(rect$x, function(z) dmixnorm(z, 0, 0, 1, 1, t))-rect$y, type = c("2"))
  }
}

y1 = rep(0, length(bw))
y2 = rep(0, length(bw))
y3 = rep(0, length(bw))

for (k in 1:length(bw)){
  y1[k] = mean(gau_err[,k])
  y2[k] = mean(epane_err[,k])
  y3[k] = mean(rect_err[,k])
}
```

```{r kernelerror, fig.cap="The graphs captures the 2-Norm error of different Kernels to its theoretical pdf across a variety of different bandwidths"}
par(mfrow=c(1,3))
plot(bw, y1, col = "red", type = "l", xlab = "Bandwidth", ylab = "Err Gaussian Kernel")
plot(bw, y2, col = "blue", type = "l", xlab = "Bandwidth", ylab = "Err Epanechnikov Kernel")
plot(bw, y3, col = "green", type = "l", xlab = "Bandwidth", ylab = "Err Rectangular Kernel")
mtext("Error plots of Kernels to Density", side = 3, line = - 2, outer = TRUE)
```

# Conlcusion
Figure \@ref(fig:kernelerror) demonstrates pretty clearly that for each choosen bandwidth different Kernel estimators, given by the Gaussian Kernel, the Epanechnikov and the Rectangular Kernel are nearly identical. Therefore the error given by taking the norm of the difference between the Kernel and its theoretical pdf has a stronger dependence on the bandwidth than the Kernel.
